Low prevalence of \emph{a priori} power analyses in motor behavior research

### 2022-10-19 docs/revision-1/r1_manuscript.tex
init

\abstract{%
\emph{A priori} power analyses can ensure studies are unlikely to miss interesting effects. Recent metascience has suggested that kinesiology research may be underpowered and selectively reported. Here, we examined whether power analyses are being used to ensure informative studies in motor behavior. We reviewed every article published in three motor behavior journals between January 2019 and June 2021. Power analyses were reported in 13\% of studies (\emph{k} = 636) that tested a hypothesis. No study targeted the smallest effect size of interest. Most studies with a power analysis relied on estimates from previous experiments, pilot studies, or benchmarks to determine the effect size of interest. Studies without a power analysis reported support for their main hypothesis 85\% of the time, while studies with a power analysis found support 76\% of the time. The median sample sizes were \emph{n} = 17.5 without a power analysis and \emph{n} = 16 with a power analysis, suggesting the typical study design was underpowered for all but the largest plausible effect size. At present, power analyses are not being used to optimize the informativeness of motor behavior research. Adoption of this widely recommended practice may greatly enhance the credibility of the motor behavior literature.
}

@@

Standardized effect sizes also have conventional benchmarks that researchers may rely on when designing studies. Recent metascience suggests doing so is likely to result in underpowered research designs in practice (Lovakov \& Agadullina, 2021). Instead of relying on benchmarks, some researchers may base their effect size target on a previous study or the results of a pilot study. However, large multi-lab replication studies have revealed that original studies may overestimate the true effect of an independent variable by 100\% to 400\% (Klein et al., 2018; Open Science Collaboration, 2015). Pilot studies are often even less helpful, as they tend to be smaller than published experiments so their estimates are even more imprecise (Albers \& Lakens, 2018; Kraemer et al., 2006; Lakens \& Evers, 2014). When available, meta-analyses provide an effect size estimate based on the aggregation of available data. However, selective reporting of results can distort meta-analytic estimates and it can be difficult to correct for reporting bias (Carter et al., 2019; Thornton \& Lee, 2000). Nevertheless, estimates that have been corrected for reporting bias are more accurate than naïve random effects estimates and should be used when available (Carter et al., 2019).

@@

Our design and analysis plan was preregistered after piloting our methods on a subsample of 40 papers. The preregistration, data, and code for this study is available using either of the following links: \url{https://osf.io/wsdpv/} \censor{or} \censor{$https://github.com/cartermaclab/proj_power-motor-behaviour$}.

@@

Calculating \emph{a priori} power for this study required estimating the final sample size and proportion of studies that would include a power analysis. Based on the pilot sample of articles, we estimated that the total number of studies we would extract would be approximately 500. The actual number was 636. We reasoned that if 10\% of those studies included a power analysis, we would have 50 studies with power analyses and 450 studies without power analyses in our sample. The actual numbers were 13\%, 85 and 551. Based on our rough estimates, we conducted simulations to estimate our power to detect differences in positive result rates of various plausible sizes. We based our expected positive results rate in studies without power analyses on estimates for psychology overall at 91.5\% (Fanelli, 2010). We observed that, if our estimates were accurate, we would have 90\% power to detect a difference of 16.5\%, or a positive result rate of 75\% in experiments with power analyses. Similarly, we estimated we would have 80\% power to identify a positive result rate of 77.7\% as significantly different. Unfortunately, if our estimated group sizes were accurate, we would have had low power (32\%) to detect our smallest effect size of interest (6\%). Given the actual sample sizes we observed, we had even greater power than planned to observe the effects we considered.

@@

All articles published in the \emph{Journal of Motor Learning and Development}, \emph{Human Movement Science}, and the \emph{Journal of Motor Behavior} between January 2019 to June 2021 were uploaded to Covidence systematic review software and screened for inclusion (Figure \ref{fig:fig1}). In total, 704 articles were reviewed. To be included in the analysis, studies were required to meet the following criteria: a) must be a primary study, b) must test a hypothesis, including the null hypothesis, c) there must be sufficient information available to adequately evaluate the criteria, and d) we must have access to the full-text. From the original 704 articles, 607 articles included at least one study that met the inclusion criteria and were included in the final analysis. Ninety-seven articles were excluded from the analysis for the following reasons: a) the studies were not primarily quantitative (63 studies), b) the studies made no hypothesis (27 studies), or there was insufficient information or a faulty DOI to assess the paper (7). The 607 included articles contributed a total of 636 eligible studies to the analysis.

@@

\hypertarget{statistical-analysis}{%
\subsection{Statistical Analysis}\label{statistical-analysis}}

@@

We calculated the percentage of studies that conducted a power analysis with a) each effect justification and b) each power target. A two-sided, six sample proportion test was conducted to test whether at least two different effect size justifications in power analyses led to different rates of positive results. A two-sided, 11-sample proportion test was conducted to test whether at least two power targets resulted in a different rate of positive results. Lastly, we conducted a two-tailed Welch's \emph{t}-test to determine whether studies with power analyses had different sample sizes compared to studies without power analyses. Given the data were highly skewed, we also conducted a sensitivity analysis using a shift function (Rousselet et al., 2017; Rousselet \& Wilcox, 2020; Wilcox, 2021).

@@

\hypertarget{proportion-of-studies-with-a-power-analysis}{%
\subsection{Proportion of Studies with a Power Analysis}\label{proportion-of-studies-with-a-power-analysis}}

Out of 636 total studies, 85 included a power analysis and 551 did not. Therefore, 13\% of all studies sampled reported the results of a power analysis.

\hypertarget{difference-in-positivity-rates-between-studies-with-and-without-a-power-analysis}{%
\subsection{Difference in Positivity Rates between Studies with and without a Power Analysis}\label{difference-in-positivity-rates-between-studies-with-and-without-a-power-analysis}}

As shown in Figure \ref{fig:fig2}, studies that did not include a power analysis reported finding support for their primary hypothesis 85\% of the time (95\% CI {[}82\%, 88\%{]}), while studies that included a power analysis found support 76\% of the time (95\% CI {[}66\%, 85\%{]}). The difference in positivity rates was not statistically significant, \(\chi^2\)= 3.47, \emph{df} = 1, \emph{p} = .06. The difference is positivity rates was not significantly smaller than our smallest effect size of interest, \emph{Z} = .546, \emph{p} = .71.

@@

\hypertarget{justifications-for-effect-sizes-used-in-power-analyses}{%
\subsection{Justifications for Effect Sizes used in Power Analyses}\label{justifications-for-effect-sizes-used-in-power-analyses}}

The most common justification reported in our sample was to base the expected effect size on a previous study (\emph{n} = 37), accounting for 44\% of all justifications. The second most common justification was to provide no justification at all (\emph{n} = 20), which occurred in 24\% of studies that included a power analysis. Cohen's benchmarks for small, medium, and large effects (\emph{n} = 19) were used in 22\% of studies. Pilot studies (\emph{n} = 9) were used as justification in 11\% of the sample.

\hypertarget{power-levels-targeted-in-power-analyses}{%
\subsection{Power Levels Targeted in Power Analyses}\label{power-levels-targeted-in-power-analyses}}

The most frequently targeted power was 80\%, which was chosen in 65\% of studies with a power analysis (\emph{n} = 55). The next most common power target was 95\%, accounting for 14\% of all power targets (\emph{n} = 12); followed by 90\% power, occurring in 11\% of power analyses (\emph{n} = 9). Two studies did not state their targeted power and several idiosyncratic power targets (96.7\%, 96\%, 95.33\%, 85\%, 75\%, 70\%, and 20\%) were reported only once.

\hypertarget{difference-in-positivity-rates-as-a-function-of-effect-size-justification}{%
\subsection{Difference in Positivity Rates as a Function of Effect Size Justification}\label{difference-in-positivity-rates-as-a-function-of-effect-size-justification}}

Figure \ref{fig:fig3} illustrates the proportion of positive results for the four effect size justifications we found in our sample. Positivity rates were 100\% for pilot study justification (9/9), 90\% for studies with no justification (18/20), 68\% for studies based on benchmarks (13/19), and 68\% for studies based on previous studies (25/37). There was no significant difference between the positivity rates of any two justifications, \(\chi^2\) = 7.12, \emph{df} = 3, \emph{p} = .068.

@@

\hypertarget{difference-in-positivity-rates-as-a-function-of-target-power}{%
\subsection{Difference in Positivity Rates as a Function of Target Power}\label{difference-in-positivity-rates-as-a-function-of-target-power}}

Studies that targeted 80\% power found support for their hypotheses 68\% of the time (38/56). Studies that aimed for 90\% power found support 100\% of the time (8/8) and studies that aimed for 95\% power found support 75\% of the time (9/12). All studies that set an idiosyncratic power target or no target at all found support for their hypotheses (10/10). There was no significant difference between target power values, \(\chi^2\)= 7.22, \emph{df} = 10, \emph{p} = .70.

\hypertarget{difference-in-sample-size-between-studies-with-and-without-a-power-analysis}{%
\subsection{Difference in Sample Size between Studies with and without a Power Analysis}\label{difference-in-sample-size-between-studies-with-and-without-a-power-analysis}}

Studies that included a power analysis had significantly smaller mean sample sizes (\emph{M} = 21.91) than studies that did not include a power analysis (\emph{M} = 40.98), \emph{t}(624.59)= 3.43, \emph{p} = .001. However, sample sizes were highly skewed---especially among studies without a power analysis. The median sample for those studies (\emph{Mdn} = 17.5) was similar to the sample sizes of studies with a power analysis (\emph{Mdn} = 16). We conducted a shift function as a sensitivity analysis and the results indicated no significant difference in sample size between studies with and without power analyses at any decile of their distributions.

@@

The purpose of this study was to investigate how frequently power analyses are reported in motor behavior articles, the justifications used for the effect size estimates, and the relationship between reporting power analyses and reporting positive results. We reviewed every article published in the \emph{Journal of Motor Behavior}, the \emph{Journal of Motor Learning and Development}, and \emph{Human Movement Science} between January 2019 and June 2021 and identified 636 studies that tested a hypothesis. Of those 636 studies, 85 of them included a power analysis (13\%). The rate of positive results was 85\% overall and 76\% when a power analysis was reported. The positive result rate was not significantly different between various effect size justifications or power targets.

Our results suggest that motor behavior research has not yet widely adopted power analyses to inform study design. When power analyses were reported, we observed a range of suboptimal effect size justifications. For example, 63\% of studies that reported a power analysis based their effect size assumption on a previous study, a pilot study, or on effect size benchmarks. Another 24\% of studies provided no justification at all. Each of these justifications (or lack thereof) are undesirable for different reasons. Previous studies---and especially pilot studies---are likely to provide exaggerated or noisy estimates of the unknown true effect. Effect size benchmarks may not match well the typical effect sizes one may find in their respective research area (Lovakov \& Agadullina, 2021). Further, Cohen's (Cohen, 1988) benchmarks differ depending on which analysis is used in a power analysis. A medium effect is over twice as large for a multiple regression analysis as compared to a \emph{t}-test (see Correll et al., 2020 for a discussion with additional examples). Not one study in the sample performed a power analysis based on their smallest effect size of interest. Power analyses can be an effective tool for researchers to ensure their studies are not underpowered, but to do so the smallest effects of interest need to be targeted.

The rate of positive results observed in this study suggest that positive findings are overrepresented in the motor behavior literature. While the studies in our sample reported positive results 84\% of the time, the median per group sample size was \textasciitilde17, which would provide 84\% power to detect \emph{d} = 1.05 with an independent \emph{t}-test or \emph{d} = .76 with a dependent \emph{t}-test. In comparison, the most optimistic estimates for well-known motor behavior phenomena are much smaller. For example, the effect of feedback frequency on motor performance (\emph{d} = .19, McKay et al., 2022), self-controlled practice on retention performance (\emph{d} = .54, McKay et al., in-press), enhanced expectancies on retention performance (\emph{d} = .54, Bacelar et al., in-press), and external focus of attention on retention performance (\emph{d} = .58, Chua et al., 2021). Estimates for the true effects of these phenomena that have been corrected for reporting-bias are markedly smaller, ranging from \emph{d} = 0 to \emph{d} = .25. Assuming the average effects investigated by the studies in our sample were similar to the optimistic estimates for other motor behavior effects, it is likely this literature was underpowered on average and potentially heavily censored.

All three journals that we sampled from either explicitly mention power in their instructions for authors (\emph{Human Movement Science}), or reference JARS (\emph{Journal of Motor Learning and Development}) or CONSORT (\emph{Journal of Motor Behavior}) reporting standards; both of which include power analyses. Therefore, adoption of power analyses targeting interesting effects does not require a policy shift, simply the enforcement of current guidelines. Since no studies in this sample powered for the smallest effect size of interest, if \emph{Human Movement Science}, the \emph{Journal of Motor Behavior}, and/or the \emph{Journal of Motor Learning and Development} enforce their existing guidelines then their future publications will look dramatically different. We believe this is a promising path forward to increase the reliability of motor behavior research and the evidence-based recommendations for coaching and rehabilitation.

@@

Our results suggest that power analyses targeting the smallest effect size of interest (Lakens, 2022a) have the potential to change the state of the motor behavior literature. Hypothesis tests are the norm in this space, yet power calculations targeting interesting effects are not. It is logical for researchers to plan studies with a high probability of producing informative results and it is consistent with current reporting standards (Appelbaum et al., 2018). Given the recent concern about the reliability of established motor behavior phenomena Mesquida et al. (2022), we believe power analyses have an important role to play in increasing the credibility of our field.

@@

@@

>>>>

\abstract{%
\emph{A priori} power analyses can ensure studies are unlikely to miss interesting effects. Recent metascience has suggested that kinesiology research may be underpowered and selectively reported. Here, we examined whether power analyses are being used to ensure informative studies in motor behavior. We reviewed every article published in three motor behavior journals between January 2019 and June 2021. Power analyses were reported in 13\% of studies (\emph{k} = \textcolor{blue}{635}) that tested a hypothesis. No study targeted the smallest effect size of interest. Most studies with a power analysis relied on estimates from previous experiments, pilot studies, or benchmarks to determine the effect size of interest. Studies without a power analysis reported support for their main hypothesis 85\% of the time, while studies with a power analysis found support \textcolor{blue}{77}\% of the time. The median sample sizes were \emph{n} = 17.5 without a power analysis and \emph{n} = 16 with a power analysis, suggesting the typical study design was underpowered for all but the largest plausible effect size. At present, power analyses are not being used to optimize the informativeness of motor behavior research. Adoption of this widely recommended practice may greatly enhance the credibility of the motor behavior literature.
}

@@

Standardized effect sizes also have conventional benchmarks that researchers may rely on when designing studies. Recent metascience suggests doing so is likely to result in underpowered research designs in practice (Lovakov \& Agadullina, 2021). Instead of relying on benchmarks, some researchers may base their effect size target on a previous study or the results of a pilot study. However, large multi-lab replication studies have revealed that original studies may overestimate the true effect of an independent variable by 100\% to 400\% (Klein et al., 2018; Open Science Collaboration, 2015). Pilot studies are often even less helpful, as they tend to be smaller than published experiments so their estimates are even more imprecise (Albers \& Lakens, 2018; Kraemer et al., 2006; Lakens \& Evers, 2014). When available, meta-analyses provide an effect size estimate based on the aggregation of available data. However, selective reporting of results can distort meta-analytic estimates and it can be difficult to correct for reporting bias (Carter et al., 2019; Thornton \& Lee, 2000). Nevertheless, estimates that have been corrected for reporting bias are more accurate than naïve random effects estimates and should be used when available (Carter et al., 2019). \textcolor{blue}{Targeting the estimated mean effect will only provide the desired power if the effect is average or above average in the specific context being investigated. Therefore, even if researchers are privy to a precise and unbiased estimate of the mean effect they are studying, they still might choose a different effect to target in their power analysis.}

@@

Our design and analysis plan was preregistered after piloting our methods on a subsample of 40 papers. The preregistration, materials, data, and code are available using either of the following links: \url{https://osf.io/wsdpv/} \censor{or} \censor{$https://github.com/cartermaclab/proj_power-motor-behaviour$}.

@@

Calculating \emph{a priori} power for this study required estimating the final sample size and proportion of studies that would include a power analysis. Based on the pilot sample of articles, we estimated that the total number of studies we would extract would be approximately 500. The actual number was \textcolor{blue}{635}. We reasoned that if 10\% of those studies included a power analysis, we would have 50 studies with power analyses and 450 studies without power analyses in our sample. The actual numbers were 13\%, \textcolor{blue}{84} and 551. Based on our rough estimates, we conducted simulations to estimate our power to detect differences in positive result rates of various plausible sizes. We based our expected positive results rate in studies without power analyses on estimates for psychology overall at 91.5\% (Fanelli, 2010). We observed that, if our estimates were accurate, we would have 90\% power to detect a difference of 16.5\%, or a positive result rate of 75\% in experiments with power analyses. Similarly, we estimated we would have 80\% power to identify a positive result rate of 77.7\% as significantly different. Unfortunately, if our estimated group sizes were accurate, we would have had low power (32\%) to detect our smallest effect size of interest (6\%). Given the actual sample sizes we observed, we had even greater power than planned to observe the effects we considered.

@@

All articles published in the \emph{Journal of Motor Learning and Development}, \emph{Human Movement Science}, and the \emph{Journal of Motor Behavior} between January 2019 to June 2021 were uploaded to Covidence systematic review software and screened for inclusion (Figure \ref{fig:fig1}). In total, 704 articles were reviewed. To be included in the analysis, studies were required to meet the following criteria: a) must be a primary study, b) must test a hypothesis, including the null hypothesis, c) there must be sufficient information available to adequately evaluate the criteria, and d) we must have access to the full-text. From the original 704 articles, \textcolor{blue}{606} articles included at least one study that met the inclusion criteria and were included in the final analysis. Ninety-\textcolor{blue}{eight} articles were excluded from the analysis for the following reasons: a) the studies were not primarily quantitative (63 studies), b) the studies made no hypothesis (27 studies), or there was insufficient information or a faulty DOI to assess the paper (\textcolor{blue}{8}). The \textcolor{blue}{606} included articles contributed a total of \textcolor{blue}{635} eligible studies to the analysis.

@@

\hypertarget{statistical-analysis}{%
\subsection{Statistical analysis}\label{statistical-analysis}}

\hypertarget{pre-registered-analyses}{%
\subsubsection{Pre-registered analyses}\label{pre-registered-analyses}}

@@

We calculated the percentage of studies that conducted a power analysis with a) each effect justification and b) each power target. A two-sided, six sample proportion test was conducted to test whether at least two different effect size justifications in power analyses led to different rates of positive results. A two-sided, 11-sample proportion test was conducted to test whether at least two power targets resulted in a different rate of positive results. We conducted a two-tailed Welch's \emph{t}-test to determine whether studies with power analyses had different sample sizes compared to studies without power analyses. Given the data were highly skewed, we also conducted a sensitivity analysis using a shift function (Rousselet et al., 2017; Rousselet \& Wilcox, 2020; Wilcox, 2021).

\hypertarget{exploratory-analysis}{%
\subsubsection{Exploratory analysis}\label{exploratory-analysis}}

\textcolor{blue}{We used a 3-sample proportion test to compare the proportion of studies including a power analysis in the three journals surveyed.}

@@

\hypertarget{pre-registered-analyses-1}{%
\subsection{Pre-registered analyses}\label{pre-registered-analyses-1}}

\hypertarget{proportion-of-studies-with-a-power-analysis}{%
\subsubsection{Proportion of studies with a power analysis}\label{proportion-of-studies-with-a-power-analysis}}

Out of \textcolor{blue}{635} total studies, \textcolor{blue}{84} included a power analysis and 551 did not\textcolor{blue}{; thus,} 13\% of all studies sampled reported the results of a power analysis.

\hypertarget{difference-in-positivity-rates-between-studies-with-and-without-a-power-analysis}{%
\subsubsection{Difference in positivity rates between studies with and without a power analysis}\label{difference-in-positivity-rates-between-studies-with-and-without-a-power-analysis}}

As shown in Figure \ref{fig:fig2}, studies that did not include a power analysis reported finding support for their primary hypothesis 85\% of the time (95\% CI {[}82\%, 88\%{]}), while studies that included a power analysis found support \textcolor{blue}{77}\% of the time (95\% CI {[}\textcolor{blue}{67}\%, \textcolor{blue}{86}\%{]}). The difference in positivity rates was not statistically significant, \(\chi^2\)= \textcolor{blue}{2.71}, \emph{df} = 1, \emph{p} = \textcolor{blue}{.10}. The difference is positivity rates was not significantly smaller than our smallest effect size of interest, \emph{Z} = \textcolor{blue}{.361}, \emph{p} = \textcolor{blue}{.641}.

@@

\hypertarget{justifications-for-effect-sizes-used-in-power-analyses}{%
\subsubsection{Justifications for effect sizes used in power analyses}\label{justifications-for-effect-sizes-used-in-power-analyses}}

The most common justification reported in our sample was to base the expected effect size on a previous study (\emph{n} = 37), accounting for 44\% of all justifications. The second most common justification was to provide no justification at all (\emph{n} = 20), which occurred in 24\% of studies that included a power analysis. Cohen's benchmarks for small, medium, and large effects (\emph{n} = \textcolor{blue}{18}) were used in \textcolor{blue}{21}\% of studies. Pilot studies (\emph{n} = 9) were used as justification in 11\% of the sample.

\hypertarget{power-levels-targeted-in-power-analyses}{%
\subsubsection{Power levels targeted in power analyses}\label{power-levels-targeted-in-power-analyses}}

The most frequently targeted power was 80\%, which was chosen in \textcolor{blue}{64}\% of studies with a power analysis (\emph{n} = \textcolor{blue}{54}). The next most common power target was 95\%, accounting for 14\% of all power targets (\emph{n} = 12); followed by 90\% power, occurring in 11\% of power analyses (\emph{n} = 9). Two studies did not state their targeted power and several idiosyncratic power targets (96.7\%, 96\%, 95.33\%, 85\%, 75\%, 70\%, and 20\%) were reported only once.

\hypertarget{difference-in-positivity-rates-as-a-function-of-effect-size-justification}{%
\subsubsection{Difference in positivity rates as a function of effect size justification}\label{difference-in-positivity-rates-as-a-function-of-effect-size-justification}}

Figure \ref{fig:fig3} illustrates the proportion of positive results for the four effect size justifications we found in our sample. Positivity rates were 100\% for pilot study justification (9/9), 90\% for studies with no justification (18/20), \textcolor{blue}{72}\% for studies based on benchmarks (13/\textcolor{blue}{18}), and 68\% for studies based on previous studies (25/37). There was no significant difference between the positivity rates of any two justifications, \(\chi^2\) = \textcolor{blue}{6.76}, \emph{df} = 3, \emph{p} = \textcolor{blue}{.08}.

@@

\hypertarget{difference-in-positivity-rates-as-a-function-of-target-power}{%
\subsubsection{Difference in positivity rates as a function of target power}\label{difference-in-positivity-rates-as-a-function-of-target-power}}

Studies that targeted 80\% power found support for their hypotheses \textcolor{blue}{70}\% of the time (38/\textcolor{blue}{54}). Studies that aimed for 90\% power found support 100\% of the time (8/8) and studies that aimed for 95\% power found support 75\% of the time (9/12). All studies that set an idiosyncratic power target or no target at all found support for their hypotheses (10/10). There was no significant difference between target power values, \(\chi^2\)= \textcolor{blue}{6.82}, \emph{df} = 10, \emph{p} = \textcolor{blue}{.743}.

\hypertarget{difference-in-sample-size-between-studies-with-and-without-a-power-analysis}{%
\subsubsection{Difference in sample size between studies with and without a power analysis}\label{difference-in-sample-size-between-studies-with-and-without-a-power-analysis}}

Studies that included a power analysis had significantly smaller mean sample sizes (\emph{M} = \textcolor{blue}{21.75}) than studies that did not include a power analysis (\emph{M} = 40.98), \emph{t}(\textcolor{blue}{624.63})= \textcolor{blue}{3.45}, \emph{p} = .001. However, sample sizes were highly skewed---especially among studies without a power analysis. The median sample for those studies (\emph{Mdn} = 17.5) was similar to the sample sizes of studies with a power analysis (\emph{Mdn} = 16). We conducted a shift function as a sensitivity analysis and the results indicated no significant difference in sample size between studies with and without power analyses at any decile of their distributions.

\hypertarget{exploratory-analysis-1}{%
\subsection{Exploratory analysis}\label{exploratory-analysis-1}}

\hypertarget{difference-among-journals-in-proportion-of-studies-with-a-power-analysis}{%
\subsubsection{Difference among journals in proportion of studies with a power analysis}\label{difference-among-journals-in-proportion-of-studies-with-a-power-analysis}}

\textcolor{blue}{Studies published in Human Movement Science included a power analysis (59/271) 18\% of the time, the Journal of Motor Learning and Development (9/128) 10\% of the time, and the Journal of Motor Behavior (16/152) 7\% of the time. The 3-samples proportions test was significant, $\chi^2$ = 13.52, \emph{df} = 2, \emph{p} = .001.}

@@

The purpose of this study was to investigate how frequently power analyses are reported in motor behavior articles, the justifications used for the effect size estimates, and the relationship between reporting power analyses and reporting positive results. We reviewed every article published in the \emph{Journal of Motor Behavior}, the \emph{Journal of Motor Learning and Development}, and \emph{Human Movement Science} between January 2019 and June 2021 and identified 63\textcolor{blue}{5} studies that tested a hypothesis. Of those 63\textcolor{blue}{5} studies, 8\textcolor{blue}{4} included a power analysis (13\%). The rate of positive results was 85\% overall and 7\textcolor{blue}{7}\% when a power analysis was reported. The positive result rate was not significantly different between various effect size justifications or power targets.

Our results \textcolor{blue}{cause us to wonder whether} motor behavior research has not yet widely adopted power analyses to inform study design. When power analyses were reported, we observed a range of suboptimal effect size justifications. \textcolor{blue}{Specifically, 76\%} of studies that reported a power analysis based their effect size assumption on a previous study \textcolor{blue}{(\emph{n} = 37, 44\%)}, a pilot study \textcolor{blue}{(\emph{n} = 9, 11\%)}, or on effect size benchmarks \textcolor{blue}{(\emph{n} = 18, 21\%)}. \textcolor{blue}{The remaining} 24\% of studies provided no justification at all. Each of these justifications (or lack thereof) are undesirable for different reasons. Previous studies---and especially pilot studies---are likely to provide exaggerated or noisy estimates of the unknown true effect. Effect size benchmarks may not match well the typical effect sizes one may find in their respective research area (Lovakov \& Agadullina, 2021). Further, Cohen's (Cohen, 1988) benchmarks differ depending on which analysis is used in a power analysis. A medium effect is over twice as large for a multiple regression analysis as compared to a \emph{t}-test (see Correll et al., 2020 for a discussion with additional examples). Not one study in the sample performed a power analysis based on their smallest effect size of interest. Power analyses can be an effective tool for researchers to ensure their studies are not underpowered, but to do so the smallest effects of interest need to be targeted.

The rate of positive results observed in this study suggest that positive findings are overrepresented in the motor behavior literature. While the studies in our sample reported positive results 84\% of the time, the median per group sample size was \textasciitilde17, which would provide 84\% power to detect \emph{d} = 1.05 with an independent \emph{t}-test or \emph{d} = .76 with a dependent \emph{t}-test. In comparison, the most optimistic estimates for well-known motor behavior phenomena are much smaller. For example, the effect of feedback frequency on motor performance (\emph{d} = .19, McKay et al., 2022), self-controlled practice on retention performance (\emph{d} = .54, McKay et al., in-press), enhanced expectancies on retention performance (\emph{d} = .54, Bacelar et al., in-press), and external focus of attention on retention performance (\emph{d} = .58, Chua et al., 2021). Estimates for the true effects of these phenomena that have been corrected for reporting-bias are markedly smaller, ranging from \emph{d} = 0 to \emph{d} = .25. Assuming the average effects investigated by the studies in our sample were similar to the optimistic estimates for other motor behavior effects, it is likely this literature was underpowered on average and potentially heavily censored.\footnote{\textcolor{blue}{We use the term "censored" here to mean that some experiments fail to enter the published literature for results-dependent reasons, commonly called publication bias or reporting bias.}}

All three journals that we sampled from either explicitly mention power in their instructions for authors (\emph{Human Movement Science}), or reference JARS (\emph{Journal of Motor Learning and Development}) or CONSORT (\emph{Journal of Motor Behavior}) reporting standards; both of which include power analyses. \textcolor{blue}{Perhaps the approach of more directly instructing authors to include a power analysis explains why Human Movement Science had a higher proportion of studies with a power analysis. This post-hoc explanation is consistent with the modest increase in power analysis reporting observed following adoption of guidelines requiring sample size justification in the journal Gait \& Posture} (McCrum et al., 2022). \textcolor{blue}{ However, less than one in five studies in \emph{Human Movement Science} included a power analysis so including explicit instructions to authors is not sufficient to promote widespread reporting. It seems the} adoption of power analyses targeting interesting effects does not require a policy shift, simply the enforcement of current guidelines. Since no studies in this sample powered for the smallest effect size of interest, if \emph{Human Movement Science}, the \emph{Journal of Motor Behavior}, and/or the \emph{Journal of Motor Learning and Development} enforce their existing guidelines then their future publications \textcolor{blue}{might} look dramatically different. \textcolor{blue}{Even if power analysis reporting becomes the norm, it may remain unclear whether power analyses were conducted during the design of the study or at a later stage. Therefore, the most credible confirmatory studies will have preregistered power analyses. A bold journal might even require it.} We believe this is a promising path forward to increase the reliability of motor behavior research and the evidence-based recommendations for coaching and rehabilitation.

@@

Our results suggest that power analyses targeting the smallest effect size of interest (Lakens, 2022a) have the potential to change the state of the motor behavior literature. Hypothesis tests are the norm in this space, yet power calculations targeting interesting effects are not. It is logical for researchers to plan studies with a high probability of producing informative results and it is consistent with current reporting standards (Appelbaum et al., 2018). Given the recent concern about the reliability of established motor behavior phenomena (Lohse et al., 2016; McKay et al., 2022, in-press; Mesquida et al., 2022), we believe power analyses have an important role to play in increasing the credibility of our field.

@@

\hypertarget{author-contributions-credit-taxonomy}{%
\subsection{Author Contributions (CRediT Taxonomy)}\label{author-contributions-credit-taxonomy}}

\noindent Conceptualization: \censor{BM, MJC}\\
\noindent Data curation: \censor{BM, MJC}\\
\noindent Formal analysis: \censor{BM}\\
\noindent Funding acquisition: \censor{MJC}\\
\noindent Investigation: \censor{BM, AC, MV, GJ, CT, HB, JH}\\
\noindent Methodology: \censor{BM, MJC}\\
\noindent Project administration: \censor{BM, MJC}\\
\noindent Software: \censor{BM, MJC}\\
\noindent Supervision: \censor{BM, MJC}\\
\noindent Validation: \censor{BM, MJC}\\
\noindent Visualization: \censor{BM, MJC}\\
\noindent Writing -- original draft: \censor{BM, AC, MV, GJ, CT, HB, JH, MJC}\\
\noindent Writing -- review \& editing: \censor{BM, AC, MV, GJ, CT, HB, JH, MJC}

@@

\leavevmode\vadjust pre{\hypertarget{ref-McCrum2022}{}}%
McCrum, C., Beek, J. van, Schumacher, C., Janssen, S., \& Van Hooren, B. (2022). Sample size justifications in gait \& posture. \emph{Gait \& Posture}, \emph{92}, 333--337.
